Lesson 3
========================================================

***

### What to Do First?
Notes:

***

### Pseudo-Facebook User Data
Notes:

```{r Pseudo-Facebook User Data}
getwd()
list.files()
pf <- read.csv("pseudo_facebook.tsv", sep="", header=TRUE);
```

***

### Histogram of Users' Birthdays
Notes:

#```{r Histogram of Users\' Birthdays}
library(ggplot2)
hist(user_data, dob_month)
#```

***

#### What are some things that you notice about this histogram?
Response: Most people have birthday in january! Reason: birthday not entered or unknown, so default set to january?

***

### Moira's Investigation
Notes:

***

### Estimating Your Audience Size
Notes:

***

#### Think about a time when you posted a specific message or shared a photo on Facebook. What was it?
Response:

#### How many of your friends do you think saw that post?
Response:

#### Think about what percent of your friends on Facebook see any posts or comments that you make in a month. What percent do you think that is?
Response:

***

### Perceived Audience Size
Notes:

***
### Faceting
Notes:

```{r Faceting}
ggplot(data = pf, aes(x = dob_day)) +
  geom_histogram() +
  scale_x_discrete(breaks = 1:31) +
  facet_wrap(~dob_month)
```

#### Let’s take another look at our plot. What stands out to you here?
Response:

***

### Be Skeptical - Outliers and Anomalies
Notes:

***

### Moira's Outlier
Notes:
#### Which case do you think applies to Moira’s outlier?
Response:

***

### Friend Count
Notes:

#### What code would you enter to create a histogram of friend counts?

```{r Friend Count}
#hist(user_data$friend_count)
ggplot(data = pf, aes(x = friend_count)) +
geom_histogram()

```

#### How is this plot similar to Moira's first plot?
Response:

***

### Limiting the Axes
Notes:

```{r Limiting the Axes}
hist(user_data$friend_count, xlim=c(0,1000))
```

### Exploring with Bin Width
Notes:

***

### Adjusting the Bin Width
Notes:

### Faceting Friend Count
```{r Faceting Friend Count}
# What code would you add to create a facet the histogram by gender?
# Add it to the code below.
qplot(x = friend_count, data = pf, binwidth = 10, facets = ~gender) +
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50))
```

***

### Omitting NA Values
Notes:

```{r Omitting NA Values}
qplot(x = friend_count, data = na.omit(pf), binwidth = 10, facets = ~gender) +
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50))
```

***

### Statistics 'by' Gender
Notes:

```{r Statistics \'by\' Gender}
by(pf$friend_count, pf$gender, summary)
```

#### Who on average has more friends: men or women?
Response: woman

#### What's the difference between the median friend count for women and men?
Response: 22

#### Why would the median be a better measure than the mean?
Response: the mean is more sensitive to outliers

***

### Tenure
Notes:

```{r Tenure}
ggplot(aes(x = tenure), data = pf) +
   geom_histogram(binwidth = 30, color = 'black', fill = '#099DD9')
```

***

#### How would you create a histogram of tenure by year?

```{r Tenure Histogram by Year}
  ggplot(aes(x = tenure / 365), data = pf) +
  geom_histogram(color = 'black', fill = '#F79420') +
  scale_x_continuous(breaks = seq(1, 7, 1), limits = c(0, 7)) +
  xlab('Number of years using Facebook') +
  ylab('Number of users in sample')
```

***

### Labeling Plots
Notes:

```{r Labeling Plots}

```

***

### User Ages
Notes:

```{r User Ages}
  ggplot(aes(x = age), data = pf) +
  geom_histogram(binwidth = 10, color = 'black', fill = '#F79420') +
  scale_x_continuous(breaks = seq(1, 7, 1), limits = c(8, 100)) +
  xlab('Age of users') +
  ylab('Number of users in sample')
```

#### What do you notice?
Response:

***

### The Spread of Memes
Notes:

***

### Lada's Money Bag Meme
Notes:

***

### Transforming Data
Notes:

```{r Friend Count}
  library(gridExtra)
  plot1 <- ggplot(aes(x = friend_count), data = pf) + geom_histogram()
  #geom_histogram(binwidth = 1, color = 'black', fill = '#F79420')
  plot2 <- plot1 + scale_x_log10();
  plot3 <- plot1 + scale_x_sqrt();
  #plot2 <- ggplot(aes(x = log10(age + 1)), data = pf) +
  #geom_histogram(binwidth = 0.1, color = 'black', fill = '#F79420')
  #plot3 <- ggplot(aes(x = sqrt(age)), data = pf) +
  #geom_histogram(binwidth = 1, color = 'black', fill = '#F79420')

  grid.arrange(plot1, plot2, plot3, ncol=1, nrow = 3)
```

***

### Add a Scaling Layer
Notes:

```{r Add a Scaling Layer}

```

***


### Frequency Polygons

```{r Frequency Polygons}
  ggplot(aes(x = www_likes, color = gender), data = pf) + geom_freqpoly(binwidth=1) + scale_x_log10()
```

***

### Likes on the Web
Notes:

```{r Likes on the Web}
by(pf$www_likes, pf$gender, sum)
```


***

### Box Plots
Notes:

```{r Box Plots}
qplot(x = gender, y =friend_count, data = subset(pf, !is.na(gender)), geom="boxplot")
```

#### Adjust the code to focus on users who have friend counts between 0 and 1000.

```{r}
qplot(x = gender, y =friend_count, data = subset(pf, !is.na(gender)), geom="boxplot") + ylim(0,1000)
qplot(x = gender, y =friend_count, data = subset(pf, !is.na(gender)), geom="boxplot") + coord_cartesian(ylim= c(0,1000))
```

***

### Box Plots, Quartiles, and Friendships
Notes:

```{r Box Plots, Quartiles, and Friendships}
by(pf$friendships_initiated, pf$gender. summary)
qplot(x = gender, y =friendships_initiated, data = subset(pf, !is.na(gender)), geom="boxplot") + coord_cartesian(ylim= c(0,100))
```

#### On average, who initiated more friendships in our sample: men or women?
Response:
#### Write about some ways that you can verify your answer.
Response:
```{r Friend Requests by Gender}
  Woman initiate more friendships, because the mean, median and 1st and 3rd quartile are higher.
```

Response:

***

### Getting Logical
Notes:

```{r Getting Logical}
pf$mobile_check_in <- NA
pf$mobile_check_in <- ifelse(pf$mobile_likes > 0, 1, 0)
pf$mobile_check_in <- factor(pf$mobile_check_in)
summary(pf$mobile_check_in)
sum(pf$mobile_check_in ==1) / length(pf$mobile_check_in)
summary(pf$mobile_check_in)[2] / (summary(pf$mobile_check_in)[1] + summary(pf$mobile_check_in)[2]) *100
```

Response:

***

### Analyzing One Variable
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!