Lesson 4
========================================================

***

### Scatterplots and Perceived Audience Size
Notes:

***

### Scatterplots
Notes:

```{r Scatterplots}
pf <- read.csv('pseudo_facebook.tsv', sep = '\t')
qplot(x = age, y = friend_count, data = pf)
```

***

#### What are some things that you notice right away?
Response: Young friends have a lot of friends!

***

### ggplot Syntax
Notes:

```{r ggplot Syntax}
ggplot(aes(x = age, y = friend_count), data = pf) + geom_point() + xlim(13,90)
```

***

### Overplotting
Notes:

```{r Overplotting}
ggplot(aes(x = age, y = friend_count), data = pf) + geom_jitter(alpha=1/20) + xlim(13,90)
```

#### What do you notice in the plot?
Response: people below 25 have a lof of friends!

***

### Coord_trans()
Notes:

```{r Coord_trans()}
ggplot(aes(x = age, y = friend_count), data = pf) + geom_point(alpha=1/20) + xlim(13,90) + coord_trans(y = "sqrt")
```

#### Look up the documentation for coord_trans() and add a layer to the plot that transforms friend_count using the square root function. Create your plot!

```{r}
?coord_trans
```

#### What do you notice?
Much better readable


***

### Alpha and Jitter
Notes:

```{r Alpha and Jitter}
ggplot(aes(x = age, y = friendships_initiated), data = pf) + geom_jitter() +  scale_x_continuous(breaks = seq(13, 120, 10)) + coord_trans(y = "sqrt")
```

***

### Overplotting and Domain Knowledge
Notes:

***

### Conditional Means
Notes:

```{r Conditional Means}
library(dplyr)
pf.fc_by_age <- pf %>% 
  group_by(age) %>%
  summarise(friend_count_mean  = mean(friend_count),
              friend_count_median = median(friend_count),
              n = n()) %>%
  arrange(age)
head(pf.fc_by_age)  
```

Create your plot!

```{r Conditional Means Plot}
ggplot(aes(x = age, y = friend_count_mean), data = pf.fc_by_age) + geom_line()
```

***

### Overlaying Summaries with Raw Data
Notes:

```{r Overlaying Summaries with Raw Data}
ggplot(aes(x = age, y = friend_count), data = pf) + 
  geom_point(alpha=1/20, position = position_jitter(h=0), color='orange') +
  coord_cartesian(xlim = c(13, 90)) + 
  coord_trans(y = "sqrt") +
  geom_line(stat = 'summary', fun.y = mean) +
  geom_line(stat = 'summary', fun.y = quantile, probs = 0.1, linetype = 2, color = 'blue')

```

#### What are some of your observations of the plot?
Response:

***

### Moira: Histogram Summary and Scatterplot
See the Instructor Notes of this video to download Moira's paper on perceived audience size and to see the final plot.

Notes:

***

### Correlation
Notes:

```{r Correlation}
?cor.test
with(data=pf, cor.test(age, friend_count))
```

Look up the documentation for the cor.test function.

What's the correlation between age and friend count? Round to three decimal places.
Response:

***

### Correlation on Subsets
Notes:

```{r Correlation on Subsets}
with(data=pf[pf$age <= 70,], cor.test(age, friend_count))
with(subset(pf, age <= 70), cor.test(age, friend_count))
```

***

### Correlation Methods
Notes:

***

## Create Scatterplots
Notes:

```{r}
ggplot(aes(x = www_likes_received, y = likes_received), data = pf) + geom_point(alpha=1/20) + coord_cartesian(xlim = c(0, 1000), ylim=c(0, 3000)) +
  coord_trans(y = "sqrt")
```

***

### Strong Correlations
Notes:

```{r Strong Correlations}
ggplot(aes(x = www_likes_received, y = likes_received), data = pf) + 
  geom_point(alpha=1/20) + 
  xlim(0, quantile(pf$www_likes_received, 0.95)) + 
  ylim(0, quantile(pf$likes_received, 0.95)) +
  geom_smooth(method = "lm", color = "red") + 
  coord_trans(y = "sqrt")
```

What's the correlation betwen the two variables? Include the top 5% of values for the variable in the calculation and round to 3 decimal places.

```{r Correlation Calcuation}
with(pf, cor.test(www_likes_received, likes_received))
```

Response:

***

### Moira on Correlation
Notes:

***

### More Caution with Correlation
Notes:

```{r More Caution With Correlation}
install.packages('alr3')
library(alr3)
?Mitchell
```

Create your plot!

```{r Temp vs Month}
ggplot(aes(x = Month, y = Temp), data = Mitchell) + 
  geom_point() 
```

***

### Noisy Scatterplots
a. Take a guess for the correlation coefficient for the scatterplot.

b. What is the actual correlation of the two variables?
(Round to the thousandths place)

```{r Noisy Scatterplots}
with(Mitchell, cor.test(Month, Temp))
```

***

### Making Sense of Data
Notes:

```{r Making Sense of Data}
range(Mitchell$Month)
ggplot(aes(x = Month, y = Temp), data = Mitchell) + 
  geom_point() +
  scale_x_discrete(breaks = seq(0, 203, 12))
```

***

### A New Perspective

What do you notice?
Response:

Watch the solution video and check out the Instructor Notes!
Notes:

***

### Understanding Noise: Age to Age Months
Notes:

```{r Understanding Noise: Age to Age Months}
ggplot(aes(x=(Month%%12),y=Temp),data=Mitchell)+
  geom_point()
library(energy)
x <- seq(0, 4*pi, pi/20)
y <- cos(x)
qplot(x = x, y = y)
dcor.ttest(x, y)
```

***

### Age with Months Means

```{r Age with Months Means}
pf$age_with_months <- (2013 - pf$dob_year) + (12 - pf$dob_month) * (1/12)
pf.fc_by_age_months <- pf %>% 
  group_by(age_with_months) %>%
  summarise(friend_count_mean  = mean(friend_count),
              friend_count_median = median(friend_count),
              n = n()) %>%
  arrange(age_with_months)
head(pf.fc_by_age_months) 

```

Programming Assignment
```{r Programming Assignment}

```

***

### Noise in Conditional Means

```{r Noise in Conditional Means}
ggplot(aes(x=age_with_months, y=friend_count_mean), data=subset(pf.fc_by_age_months, age_with_months < (71*12)))+
  geom_line() + xlim(13,70) + ylim(0,500)
```

***

### Smoothing Conditional Means
Notes:

```{r Smoothing Conditional Means}

```

***

### Which Plot to Choose?
Notes:

***

### Analyzing Two Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

