Lesson 5
========================================================

### Multivariate Data
Notes:

***

### Moira Perceived Audience Size Colored by Age
Notes:

***

### Third Qualitative Variable
Notes:

```{r Third Qualitative Variable}
ggplot(aes(x = gender, y = age),
       data = subset(pf, !is.na(gender))) + geom_histogram()

pf <- read.delim('../lesson3/pseudo_facebook.tsv')
suppressMessages(library(dplyr))

pf.fc_by_age_gender <- pf %>%
  filter(!is.na(gender)) %>%
  group_by(age, gender) %>%
  summarise(mean_friend_count  = mean(friend_count),
            median_friend_count = median(friend_count),
            n = n()) %>%
  ungroup() %>%
  arrange(age, gender) 
head(pf.fc_by_age_gender) 
```

***

### Plotting Conditional Summaries
Notes:

```{r Plotting Conditional Summaries}
ggplot(aes(x = age, y = median_friend_count, color=gender), data = pf.fc_by_age_gender) + 
  geom_line()
```

***

### Thinking in Ratios
Notes:

***

### Wide and Long Format
Notes:

***

### Reshaping Data
Notes:

```{r}
install.packages('reshape2')
library(reshape2)
pf.fc_by_age_gender.wide <- dcast(pf.fc_by_age_gender, 
                                  age ~ gender,
                                  value.var = 'median_friend_count')
head(pf.fc_by_age_gender.wide)

#OR
pf.fc_by_age_gender.wide <-     subset(pf.fc_by_age_gender[c('age', 'gender', 'median_friend_count')],                   !is.na(gender)) %>%
    spread(gender, median_friend_count) %>%
    mutate(ratio = male / female)

head(pf.fc_by_age_gender.wide)
```


***

### Ratio Plot
Notes:

```{r Ratio Plot}
ggplot(aes(x = age, y = female/male), data = pf.fc_by_age_gender.wide) + 
  geom_line() +
  geom_hline(y=1, linetype=3)
```

***

### Third Quantitative Variable
Notes:

```{r Third Quantitative Variable}
pf$year_joined <- floor(2014 - (pf$tenure / 365))
```

***

### Cut a Variable
Notes:

```{r Cut a Variable}
pf$year_joined.bucket <- cut(pf$year_joined, breaks=c(2004, 2009, 2011, 2012, 2014))
```

***

### Plotting it All Together
Notes:

```{r Plotting it All Together}
ggplot(aes(x = age, y = friend_count), data = subset(pf, !is.na(gender))) +
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = median)
```

***

### Plot the Grand Mean
Notes:

```{r Plot the Grand Mean}
ggplot(aes(x = age, y = friend_count), data = subset(pf, !is.na(gender))) +
      geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = mean) +
      geom_line(stat = 'summary', fun.y = mean, linetype=2, color = 'black') 
```

***

### Friending Rate
Notes:

```{r Friending Rate}

pf_tenure_gt_zero <- subset(pf, tenure > 0)
pf_tenure_gt_zero$friendRate = pf_tenure_gt_zero$friend_count / pf_tenure_gt_zero$tenure
median(pf_tenure_gt_zero$friendRate)
max(pf_tenure_gt_zero$friendRate)
```

***

### Friendships Initiated
Notes:

What is the median friend rate?

What is the maximum friend rate?

```{r Friendships Initiated}
pf_tenure_gt_zero <- subset(pf, tenure >= 1)
ggplot(aes(x = tenure, y = friendships_initiated / tenure), data = pf_tenure_gt_zero) +
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = median)

```

***

### Bias-Variance Tradeoff Revisited
Notes:

```{r Bias-Variance Tradeoff Revisited}

ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_line(aes(color = year_joined.bucket),
            stat = 'summary',
            fun.y = mean)

ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 30 * round(tenure / 30), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 90 * round(tenure / 90), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

#smoothing
ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure), data = subset(pf, tenure > 0)) +
geom_smooth(aes(color = year_joined.bucket))

```

***

### Sean's NFL Fan Sentiment Study
Notes:

***

### Introducing the Yogurt Data Set
Notes:

***

### Histograms Revisited
Notes:

```{r Histograms Revisited}
yo <- read.csv("yogurt.csv")
str(yo)
summary(yo)
#change the id to a factor
yo$id <- factor(yo$id)
str(yo)
ggplot(data = yo, aes(x = price)) + geom_histogram(binwidth = 10) + xlab('price of yogurt') +
  ylab('count')
length(unique(yo$price))
```

***

### Number of Purchases
Notes:

```{r Number of Purchases}
yo$all.purchases = yo$strawberry + yo$blueberry	+ yo$pina.colada	+yo$plain	+ yo$mixed.berry

transform(yo, all.purchases = yo$strawberry + yo$blueberry  + yo$pina.colada	+yo$plain	+ yo$mixed.berry)
```

***

### Prices over Time
Notes:

```{r Prices over Time}
ggplot(aes(x = time, y = price), data = yo) + 
  geom_jitter(alpha=1/20)
  
```

***

### Sampling Observations
Notes:

***

### Looking at Samples of Households

```{r Looking at Sample of Households}
set.seed(1000)
sample.ids <- sample(levels(yo$id), 16)
ggplot(aes(x = time, y = price),
       data = subset(yo, id %in% sample.ids)) +
       facet_wrap(~id) +
       geom_line() +
       geom_point(aes(size = all.purchases), pch = 1)

```

***

### The Limits of Cross Sectional Data
Notes:

***

### Many Variables
Notes:

***

### Scatterplot Matrix
Notes:

***

### Even More Variables
Notes:

***

### Heat Maps
Notes:

```{r}
nci <- read.table("nci.tsv")
colnames(nci) <- c(1:64)
```

```{r}
nci.long.samp <- melt(as.matrix(nci[1:200,]))
names(nci.long.samp) <- c("gene", "case", "value")
head(nci.long.samp)

ggplot(aes(y = gene, x = case, fill = value),
  data = nci.long.samp) +
  geom_tile() +
  scale_fill_gradientn(colours = colorRampPalette(c("blue", "red"))(100))
```


***

### Analyzing Three of More Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

