---
title: "Free Trial Screener A/B test"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
##Experiment overview

At the time of this experiment, Udacity courses currently have two options on the home page: "start free trial", and "access course materials". If the student clicks "start free trial", they will be asked to enter their credit card information, and then they will be enrolled in a free trial for the paid version of the course. After 14 days, they will automatically be charged unless they cancel first. If the student clicks "access course materials", they will be able to view the videos and take the quizzes for free, but they will not receive coaching support or a verified certificate, and they will not submit their final project for feedback.

In the experiment, Udacity tested a change where if the student clicked "start free trial", they were asked how much time they had available to devote to the course. If the student indicated 5 or more hours per week, they would be taken through the checkout process as usual. If they indicated fewer than 5 hours per week, a message would appear indicating that Udacity courses usually require a greater time commitment for successful completion, and suggesting that the student might like to access the course materials for free. At this point, the student would have the option to continue enrolling in the free trial, or access the course materials for free instead. [This screenshot](https://goo.gl/haciU6) shows what the experiment looks like.

The hypothesis was that this might set clearer expectations for students upfront, thus reducing the number of frustrated students who left the free trial because they didn't have enough time—without significantly reducing the number of students to continue past the free trial and eventually complete the course. If this hypothesis held true, Udacity could improve the overall student experience and improve coaches' capacity to support students who are likely to complete the course.

The unit of diversion is a cookie, although if the student enrolls in the free trial, they are tracked by user-id from that point forward. The same user-id cannot enroll in the free trial twice. For users that do not enroll, their user-id is not tracked in the experiment, even if they were signed in when they visited the course overview page.

## Experiment design

In this section I will describe how the experiment is designed.

### Metric choice

Invariant metrics are metrics that are expected to remain (more or less) unchanged between the control and experiment group. Evaluation metrics are the metrics that are expected to change.

* Number of cookies: The number of visitors to the homepage are not expected to change, since the popup is not visible here, so this is an **invariant metric**.
* Number of user-ids: The number of users that enroll in the free trial is expected to drop (because of the popup), so this cannot be an invariant metric. But it is also not a good evaluation metric, because this number depends on the number of clicks which can fluctuate a lot in time. The rate of the number of users between the number of clicks (gross conversion) is a better metric, since it takes into account these fluctuations
* Number of clicks: The number of unique cookies to click the "Start free trial" are not expected to change, since the popup is not yet visible here, so this is an **invariant metric**.
* Click-through-probability: Since both the number of cookies and the number of clicks are not expected to change, this is also an **invariant metric**.
* Gross-conversion: The number of user-ids to enroll in the free trial is expected to drop, while the number of clicks is expected to stay the same, so this rate is expected to drop. This rate can therefore be used as an  **evaluation metric**. The practical significance boundary, which is the boundary in order to assess if the experiment is a success, is 0.01
* Retention: This is an interesting metric because it can be an indication of how motivated users are. If less user-ids enroll, but more of these will remain enrolled, this rate will increase. However, from the analysis which I have done, it follows that it would take too long to gather enough data to evaluate this metric, so we can't use this etric.
* Net conversion: The number of user-ids to remain enrolled past the 14-day boundary is expected to drop, while the number of clicks remains the same, so this rate is also expected to drop and can be used as an **evaluation metric**. The practical significance boundary is 0.0075

### Measuring Standard Deviation

To evaluate whether the analytical estimate of standard deviation is accurate and matches the empirical standard deviation, the unit of analysis and unit of diversion are compared for each evaluation metric. To calculate the standard deviation, a distribution should be known or assumed. A binomial distribution is assumed, since the success probability p can be seen as the case where a user enrolls or clicks and (1-p) the probability the user doesn't. The standard deviation for this distribution is sqrt(p*(1-p)/n), where n is the sample size.

```{r sd, echo=FALSE}
baseline <- read.csv("final_baseline.csv", header = FALSE)
df<-data.frame()
df <- rbind(df, baseline$V2)
colnames(df) <- c("pageviews_pd", "clicks_pd", "enrollments_pd", "ct_prob", "enroll_prob", "payment_enroll_prob", "payment_click_prob")

#calculate standard deviation for binomial distribution, with p the probability of success and n the sample size
calcSD <- function(p,n){
  return(sqrt((p * (1-p) / n)))
}
#sd gross conversion
sample_size <- 5000
gross_n <- sample_size * df$ct_prob
gross_sd <- calcSD(df$enroll_prob, gross_n)
#sd net conversion
net_n <- sample_size * df$ct_prob
net_sd <- calcSD(df$payment_click_prob, net_n)
#put result in df
sd <- as.data.frame(matrix(nrow = 2, ncol = 5))
colnames(sd) <- c("metric", "p", "q=(1-p)", "n", "standard deviation")
sd[1,] <- c("gross conversion", df$enroll_prob, 1-df$enroll_prob, gross_n, gross_sd)
sd[2,] <- c("net conversion", df$payment_click_prob, 1-df$payment_click_prob, net_n, net_sd)
sd
```

When the unit of analysis is equal to the unit of division, the analytical estimate is likely to match the empirical variance. For both metrics, the unit of division and unit of analysis is the cookie. So, the analytical estimate can be used for these metrics.

### Sizing
Number of Samples vs. Power
Indicate whether you will use the Bonferroni correction during your analysis phase, and give the number of pageviews you will need to power you experiment appropriately. (These should be the answers from the "Calculating Number of Pageviews" quiz.)

```{r, echo=FALSE}
samplesize <- function(alpha, beta, p, dmin) {
  z.alpha <- qnorm(1 - alpha/2)
  z.beta <- qnorm(1 - beta)
  se.null.numerator <- sqrt(2*p*(1 - p))
  se.alt.numerator <- sqrt(p*(1 - p) + (p + dmin)*(1 - (p + dmin)))
  size <- ((z.alpha*se.null.numerator + z.beta*se.alt.numerator) / dmin)^2
  return(size)
}
#online calculator: http://www.evanmiller.org/ab-testing/sample-size.html

#gross
gross_sample_size <- round(samplesize(0.05, 0.20, 0.20625, 0.01))

#net
net_sample_size <- round(samplesize(0.05, 0.20, 0.1093125, 0.0075))

#calculate pageviews by click through rate and multiply by 2 because sample size is per variation 
gross_pageviews <- gross_sample_size * (1/df$ct_prob) * 2
net_pageviews <- net_sample_size * (1/df$ct_prob) * 2
pageviews <- max(gross_pageviews, net_pageviews)
pageviews
```

Duration vs. Exposure
Indicate what fraction of traffic you would divert to this experiment and, given this, how many days you would need to run the experiment. (These should be the answers from the "Choosing Duration and Exposure" quiz.)

```{r, echo=FALSE}
days <- ceiling(max(gross_pageviews, net_pageviews) / df$pageviews_pd)
days
```

Give your reasoning for the fraction you chose to divert. How risky do you think this experiment would be for Udacity?


## Experiment analysis

###Sanity checks
For each of your invariant metrics, give the 95% confidence interval for the value you expect to observe, the actual observed value, and whether the metric passes your sanity check. (These should be the answers from the "Sanity Checks" quiz.)

For any sanity check that did not pass, explain your best guess as to what went wrong based on the day-by-day data. Do not proceed to the rest of the analysis unless all sanity checks pass.


###Result analysis
Effect Size Tests
For each of your evaluation metrics, give a 95% confidence interval around the difference between the experiment and control groups. Indicate whether each metric is statistically and practically significant. (These should be the answers from the "Effect Size Tests" quiz.)

Sign Tests
For each of your evaluation metrics, do a sign test using the day-by-day data, and report the p-value of the sign test and whether the result is statistically significant. (These should be the answers from the "Sign Tests" quiz.)

Summary
State whether you used the Bonferroni correction, and explain why or why not. If there are any discrepancies between the effect size hypothesis tests and the sign tests, describe the discrepancy and why you think it arose.

###Recommendation
Make a recommendation and briefly describe your reasoning.


##Follow-up experiment

Give a high-level description of the follow up experiment you would run, what your hypothesis would be, what metrics you would want to measure, what your unit of diversion would be, and your reasoning for these choices.







