Quality of red wine exploration by Ger Inberg
========================================================

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using
# in your analysis in this code chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk.
# This prevents the code from displaying in the knitted HTML output.
# You should set echo=FALSE for all code chunks in your file.

library(ggplot2)
library(dplyr)
library(gridExtra)
```

```{r echo=FALSE, Load_the_Data}
# Load the Data and print some basics
redwine <-read.csv("data/wineQualityReds.csv", header=TRUE, sep = ",") 
dim(redwine)
summary(redwine)
str(redwine)
```

# Univariate Plots Section
```{r echo=FALSE, Univariate_Plots}

#histograms

#fixed.acidity
ggplot(data = redwine, aes(x=fixed.acidity)) + geom_histogram()
#volatile.acidity
ggplot(data = redwine, aes(x=volatile.acidity)) + geom_histogram()
#citric.acid
ggplot(data = redwine, aes(x=citric.acid)) + geom_histogram()
#residual.sugar
ggplot(data = redwine, aes(x=residual.sugar)) + geom_histogram()
#chlorides
ggplot(data = redwine, aes(x=chlorides)) + geom_histogram()
#free.sulfur.dioxide
ggplot(data = redwine, aes(x=free.sulfur.dioxide)) + geom_histogram()
#total.sulfur.dioxide
ggplot(data = redwine, aes(x=total.sulfur.dioxide)) + geom_histogram()
#density
ggplot(data = redwine, aes(x=density)) + geom_histogram()
#pH
ggplot(data = redwine, aes(x=pH)) + geom_histogram()
#sulphates
ggplot(data = redwine, aes(x=sulphates)) + geom_histogram()
#alcohol
ggplot(data = redwine, aes(x=alcohol)) + geom_histogram()

#quality
ggplot(data = redwine, aes(x=quality)) + geom_bar(binwidth=1)
```

# Univariate Analysis

### What is the structure of your dataset?

There are 1599 wines in the dataset with 13 features (X, fixed.acidity, volatile.acidity, citric.acid, residual.sugar, chlorides, free.sulfur.dioxide, total.sulfur.dioxide, density, pH, sulphates, alcohol and quality). The variables X and quality are integers, while the others are numbers.

Some other observations:
X is the wine number.
The density of wine is close to 1, so equals the density of water.
The median quality for a redwine is 6 and the max is 8.

### What is/are the main feature(s) of interest in your dataset?

The main feature in the data set is quality. I’d like to determine which features are best for predicting the quality of a red wine. I suspect a combination of the other variables can be used to build a predictive model to determine the quality.

### What other features in the dataset do you think will help support your investigation into your feature(s) of interest?

### Did you create any new variables from existing variables in the dataset?

### Of the features you investigated, were there any unusual distributions? Did you perform any operations on the data to tidy, adjust, or change the form of the data? If so, why did you do this?



# Bivariate Plots Section
```{r echo=FALSE, Bivariate_Plots}

```

# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. How did the feature(s) of interest vary with other features in the dataset?

### Did you observe any interesting relationships between the other features (not the main feature(s) of interest)?

### What was the strongest relationship you found?




# Multivariate Plots Section

```{r echo=FALSE, Multivariate_Plots}

```

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. Were there features that strengthened each other in terms of looking at your feature(s) of interest?

### Were there any interesting or surprising interactions between features?

### OPTIONAL: Did you create any models with your dataset? Discuss the strengths and limitations of your model.

------

# Final Plots and Summary

### Plot One
```{r echo=FALSE, Plot_One}

```

### Description One


### Plot Two
```{r echo=FALSE, Plot_Two}

```

### Description Two


### Plot Three
```{r echo=FALSE, Plot_Three}

```

### Description Three

------

# Reflection
