Wine quality exploration by Ger Inberg
========================================================

### Introduction

This document describes the exploration of a wine dataset and tries to find relations between features. More information about the dataset can be found in [wineQualityInfo](data/wineQualityInfo.txt).
This analysis consist of a univariate section, a bivariate section, a multivariate section and a Final plots and summary section. First, a summary is given about the used dataset.

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using
# in your analysis in this code chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk.
# This prevents the code from displaying in the knitted HTML output.
# You should set echo=FALSE for all code chunks in your file.

library(ggplot2)
library(dplyr)
library(gridExtra)
library(GGally)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, Load_the_Data}
# Load the Data, combine the two datasets and print some basics
redwine <-read.csv("data/wineQualityReds.csv", header=TRUE, sep = ",") 
whitewine <-read.csv("data/wineQualityWhites.csv", header=TRUE, sep = ",") 
redwine$color = 'red'
whitewine$color = 'white'
wine <- rbind(redwine, whitewine)
wine$color = factor(wine$color)
wine$alcohol_level <- cut(wine$alcohol, 
 breaks = c(quantile(wine$alcohol, prob =   0), 
            quantile(wine$alcohol, prob = .33), 
            quantile(wine$alcohol, prob = .66), 
            quantile(wine$alcohol, prob =   1)),
 labels = c("Low", "Medium", "High"), include.lowest=TRUE)
wine$quality_level <- cut(wine$quality, 
 breaks = c(quantile(wine$quality, prob =   0), 
            quantile(wine$quality, prob = .33), 
            quantile(wine$quality, prob = .66), 
            quantile(wine$quality, prob =   1)),
 labels = c("Low", "Medium", "High"),include.lowest=TRUE)

dim(wine)
summary(wine)
str(wine)
```

# Univariate Plots Section
```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots}

#plot histograms of every feature

#fixed.acidity
ggplot(data = wine, aes(x=fixed.acidity)) + geom_histogram()
summary(wine$fixed.acidity)
print("Acids are an important component of wine and it constists of a fixed and volatile part. This feature represents the fixed part and can be tartaric acid for example. The mean and median are both about 7 g/dm^3. There are some outliers at the right side with a maximum of almost 16 g/dm^3.")

#volatile.acidity
ggplot(data = wine, aes(x=volatile.acidity)) + geom_histogram()
summary(wine$volatile.acidity)
print("Volatile acidity is the amount of acetic acid in wine, which at too high of levels can lead to an unpleasant, vinegar taste. The mean and median are both about 0.30 g/dm^3. There are some outliers at the right side with a maximum of almost 1.58 g/dm^3.")

#citric.acid
ggplot(data = wine, aes(x=citric.acid)) + geom_histogram()
summary(wine$citric.acid)
print("Citric acid can add 'freshness' and flavor to wines. It has a 'normalish' distribution with a small peak at the left side and again some outliers at the right side, with a maximum value of 1.66 g/dm^3.")

#residual.sugar
ggplot(data = wine, aes(x=residual.sugar)) + geom_histogram()
summary(wine$residual.sugar)
print("Residual sugar is the amount of sugar remaining after fermentation stops. The distribution looks like the right side of a normal distribution, with a peak at 0 g/dm^3. The mean is quite bigger than the median, which is caused by one or more big outlier(s).")

#chlorides
ggplot(data = wine, aes(x=chlorides)) + geom_histogram()
summary(wine$chlorides)
print("Chlorides is the amount of salt in the wine. It has a 'normalish' distribution around +- 0.05 g/dm³ with some outliers at the right side.")

#free.sulfur.dioxide
ggplot(data = wine, aes(x=free.sulfur.dioxide)) + geom_histogram()
summary(wine$free.sulfur.dioxide)
print("Free sulfur dioxide is the free form of SO2. It has a 'normalish' distribution around +- 30 mg/dm^3 with some very high outliers (max=289 mg/dm^3)")

#total.sulfur.dioxide
ggplot(data = wine, aes(x=total.sulfur.dioxide)) + geom_histogram()
summary(wine$total.sulfur.dioxide)
print("The total sulfur dioxide is the amount of free and bound forms of S02. It has a 'normalish' distribution with a mean of +-116 mg/dm^3. There are some outliers with a high value.")

#density
ggplot(data = wine, aes(x=density)) + geom_histogram()
summary(wine$density)
print("The density represents the density of the wine, which is dependent on the percent alcohol and sugar content. It has a 'normalish' distribution with a mean of +- 0.995 and a max value of 1.0390 which is clearly an outlier.")

#pH
ggplot(data = wine, aes(x=pH)) + geom_histogram()
summary(wine$pH)
print("The pH features describes how acidic or basic a wine is on a scale from 0 (very acidic) to 14 (very basic). It has a 'normalish' distribution with a mean of +- 3.2 and a max value of +- 4.0 which is an outlier.")

#sulphates
ggplot(data = wine, aes(x=sulphates)) + geom_histogram()
summary(wine$sulphates)
print("The sulphates is a wine additive wich acts as an antimicrobial and antioxidant.It has a 'normalish' distribution with a mean of +- 0.53. It is clear from the plot that there are some outliers at the right side (max = 2.0).")

#alcohol
ggplot(data = wine, aes(x=alcohol)) + geom_histogram()
summary(wine$alcohol)
print("The percent alcohol content of the wine. The mean and median are a little more than 10%, with some outliers at the high percentages and a maximum of 14,9%.")

#color
ggplot(data = wine, aes(x=color)) + geom_histogram()
summary(wine$color)
print("The wine color which can either be red or wine. There are about 3 times as many white wines as red wines in this dataset.")

#quality
ggplot(data = wine, aes(x=quality)) + geom_bar(binwidth=1)
summary(wine$quality)
print("Quality is a score between 0 and 10. It has a 'normalish' distribution around 6 with a min and max of 3 and 9 respectively.")
```

# Univariate Analysis

### What is the structure of your dataset?

There are 1599 wines in the dataset with 13 features (fixed.acidity, volatile.acidity, citric.acid, residual.sugar, chlorides, free.sulfur.dioxide, total.sulfur.dioxide, density, pH, sulphates, alcohol, color and quality). The (output) variable quality is an integer, the color is a factor variable while the others are numbers.

Some other observations:
The density of wine is close to 1, so equals the density of water.
The median quality for a redwine is 6 and the max is 8.
There are a lot more observations for white wine compared to red wine.

### What is/are the main feature(s) of interest in your dataset?

The main feature in the data set is quality. I’d like to determine which features are best for predicting the quality of a red wine. I suspect a combination of the other variables can be used to build a predictive model to determine the quality.

### What other features in the dataset do you think will help support your investigation into your feature(s) of interest?

I think all of the other features, except density, can have an impact on the quality of the wine. Acidity and Alcohol could be major factors, because too much or less of these can make the wine unbalanced. 

### Did you create any new variables from existing variables in the dataset?

I have created alcohol_level and quality_level variables to be able to use them as a factor in the plots. Both have "low", "medium" and "high" levels.

### Of the features you investigated, were theres any unusual distributions? Did you perform any operations on the data to tidy, adjust, or change the form of the data? If so, why did you do this?

Most features that I plotted have a 'normalish' distribution, i.e. they have a pattern that looks like a normal distribution. This means that most of the values are of that feature are close to the mean and median count.
The plot of the residual.sugar however is different: it starts at a value of zero with the highest count and after that it only decreases. It looks like the right side of a normal distribution. 

# Bivariate Plots Section
```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots}

#Find out about correlations between features using ggcorr
ggcorr(wine[,1:13], label = TRUE, size=3, layout.exp = 1, hjust = 0.85)
print("The above plot displays the correlations between the features in the dataset. A darker color means a stronger correlation. A green color is a positive correlation, where as red is a negative correlation. The numbers in the boxes are the correlation coefficients. Below I have plotted some strong correlations.")

#alcohol vs density
ggplot(data = wine, aes(x=density, y=alcohol)) + 
  geom_point(alpha=0.2) + xlim(0.985, 1.005) +
  stat_smooth(method = "lm", col='red')
print("Alcohol and density have a strong correlation (-0.7) and this can clearly be seen in the plot. The range of density is quite small with most values between 0.99 and 1.00.")

#alcohol vs quality
ggplot(data = wine, aes(y=alcohol, x=factor(quality))) + 
  geom_jitter(alpha=0.4) + geom_boxplot(alpha=0.4) +
  stat_smooth(method = "lm", col='red')
print("Alcohol and quality have a correlation coefficient of 0.4. When looking at the interquartile ranges, we can see a general trend that it has higher alcohol percentages for a higher quality.")

#sulphates vs chlorides
ggplot(data = wine, aes(x=chlorides, y=sulphates)) + geom_point(alpha=0.2) + 
  xlim(0,0.2) + ylim(0, 1.5) + stat_smooth(method = "lm", col='red')
print("Sulphates and chlorides also have a correlation coefficient of 0.4. Though the correlation is not very strong, there is a weak to moderate relation visible.")

#total.sulfur.dioxide vs free.sulfur.dioxide
ggplot(data = wine, aes(x=free.sulfur.dioxide, y=total.sulfur.dioxide)) + 
  geom_point(alpha=0.2) +xlim(0,100) + ylim(0,250) + 
  stat_smooth(method = "lm", col='red')
print("The correlation coefficient between total.sulfur.dioxide and free.sulfur.dioxide is 0.7, so it is strong. This is not a big surprise because the total.sulfur.dioxide contains the free.sulfur.dioxide.")
```

# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. How did the feature(s) of interest vary with other features in the dataset?

I have made plots of features that did show a stronger correlation in the ggcorr output.
For the features alcohol and density there seems to be a negative trend, i.e. the alcohol % decreases with a higer density. Alcohol seems to have a positive effect on the quality.

### Did you observe any interesting relationships between the other features (not the main feature(s) of interest)?

I have looked at other relations between features and there seems to be a relationship between pH and citric.acid. When reading about acidity and ph levels, there is a logical explanation for this. The pH level of citric acid is 2.2 while wines mostly have a pH level between 2.9 and 3.9. So when more citric.acid, the pH level of the wine will drop.

### What was the strongest relationship you found?

The strongst relationship that I found is between the total.sulfur.dioxide and free.sulfur.dioxide. This is explainable, because according to the documentation the total sulfur dioxide includes the amount of free forms of SO2 (=free.sulfur.dioxide).

# Multivariate Plots Section

```{r echo=FALSE, message=FALSE, warning=FALSE, Multivariate_Plots}
#plots with 3 variables

# function to print regression for each quality level
printRegression = function(columns){
  for (i in levels(wine$quality_level)) { 
    print(i)
    wine_per_quality = wine[wine$quality_level %in% i, columns]
    print(cor(wine_per_quality)[2])
  }  
}

#alcohol vs density, color by quality_level
ggplot(data = wine, aes(x=density, y=alcohol, color=quality_level)) + 
  geom_point() + xlim(0.985, 1.005) +
   geom_smooth(method="lm", fill=NA)
printRegression(c(9,12))
print("The plot displays the alcohol vs density, colored by the quality level. The lines represent the correlation for the quality levels. As can be seen from the plot and from the coefficient: the better the quality, the stronger the correlation between the features.")

#alcohol vs chlorides, color by quality_level
ggplot(data = wine, aes(x=chlorides, y=alcohol, color=quality_level)) + 
  geom_point() + xlim(0, 0.2) +
   geom_smooth(method="lm", fill=NA)
printRegression(c(6,12))
print("The plot displays the alcohol vs chlorides, colored by the quality level. Also in this plot the correlation is stronger for a higher quality wine.")

#sulphates vs chlorides, color by quality_level
ggplot(data = wine, aes(x=chlorides, y=sulphates, color=quality_level)) + 
  geom_point() + xlim(0, 0.2) + ylim(0.2, 1.3) +
   geom_smooth(method="lm", fill=NA)
printRegression(c(6,11))
print("The plot displays the sulphates vs chlorides, colored by the quality level. ")
```

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. Were there features that strengthened each other in terms of looking at your feature(s) of interest?

When looking at the 3 plots, it seems like alcohol, sulphates and chlorides strengthened each other. The blue dots that stand for high quality wines are more dominant in the upper left corner (first 2 plots) or lower left corner (3rd plot). 

### Were there any interesting or surprising interactions between features?

When comparing differences between wine color, I found some interesting differences. The total acidity is for red wines quite higher than for white wines.

### OPTIONAL: Did you create any models with your dataset? Discuss the strengths and limitations of your model.

------

# Final Plots and Summary

### Plot One
```{r echo=FALSE, message=FALSE, warning=FALSE, Plot_One}
#histogram of quality + normal distribution
wineQualityLow <- range(wine$quality)[1]
wineQualityHigh <- range(wine$quality)[2]

ggplot(data = wine) + geom_histogram(aes(x=quality, y=..density..), 
                                     binwidth=1, colour="black", fill="white") + 
  stat_function(fun = dnorm, 
                args = list(mean = mean(wine$quality), 
                            sd = sd(wine$quality)), lwd = 2, col = 'red') +
  scale_x_continuous(breaks = seq(wineQualityLow, wineQualityHigh, 1), 
                     limits=c(3, 9)) +
  xlab('Quality') + ylab('Wine density') +
  ggtitle("Wine quality density and normal distribution")
```

### Description One

I have plotted the density of the quality of wines in a histogram plot. The histogram is overlapped with a normal distribution, that uses the mean and standard deviation from the quality variable. As can be seen, the histogram 'follows' the normal distribution quite well, it is a bit more shifted to the right.

### Plot Two
```{r echo=FALSE, message=FALSE, warning=FALSE, Plot_Two}
#alcohol vs chlorides, color by quality_level
ggplot(data = wine, aes(x=chlorides, y=alcohol, color=quality_level)) + 
  geom_point(alpha=0.2) +  geom_smooth(method="lm", fill=NA) +
  scale_x_continuous(breaks = seq(0,0.2,0.05), limits=c(0,0.2)) + 
  scale_y_continuous(breaks = seq(8,14,2), limits=c(8,14)) + 
  xlab('chlorides (g / dm^3)') + ylab('alcohol (% per volume)') +
  ggtitle("Alcohol % vs chlorides colored by quality_level")
```

### Description Two

I have plotted the alcohol of wines vs the chlorides and coloured it by the quality_level. The highest quality wines (with color blue) are mostly located at the upper left corner of the plot. This means that a high alcohol percentage and a low chlorides value are indicators for a high quality wine.

### Plot Three
```{r echo=FALSE, message=FALSE, warning=FALSE, Plot_Three}
#alcohol vs quality
ggplot(data = wine, aes(x=alcohol, y=quality)) + geom_point(alpha=0.2) + 
  stat_smooth(method = "lm", col='red') +
  scale_y_continuous(breaks = seq(wineQualityLow, wineQualityHigh, 1)) +  
  scale_x_continuous(breaks = seq(8, 14, 1), limits=c(8,14)) +
  xlab('alcohol (% by volume)') + ylab('quality') +
  facet_wrap( ~ color) +
  ggtitle("Quality vs alcohol faceted by wine color")
lm(quality ~ alcohol, data = redwine)
lm(quality ~ alcohol, data = whitewine)
```

### Description Three

I have plotted the quality of wine vs the alcohol percentage and facetted this by the wine color. Also, added a line representing the lineair model. As can be seen, the plots for red and white wine look quite similair. There seems to be a positive correlation between alcohol and quality, independent of the color. The lineair model line for red wines is a little bit steeper compared to the white wine plot.

# Reflection

The (combined) wine data set contains information on almost 6500 wines. I decided to add the red and wine datasets, because I was interested to see if the color of wine is a major factor in the quality of wine. In the boxplot of wine quality vs color, it can be seen that the quantiles are equal so color is not a major factor.
In the final plots, I have tried to visualize some interesting insights from the data. The quality histogram (plot1) follows almost a normal distribution, so most wines in the dataset are close to the mean and median value.
